test

### 2020-05-01 manuscript/response_to_reviews.Rmd

---
title: "Response to reviews"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*italics - temporary comments for points that remain to be addressed*

**bold - response**

@@

Comments to the Author

The study addresses the relationship between drought resistance (Rt) calculated from radial growth, and different traits and microenvironmental characteristics on 12 tree species from one forest in the Eastern US. Both the dataset and the study are meaningful, and the relationship found between Rt and height of biological importance. However, I think there are a number of issues that need to be worked out.

- It is important to better discuss the proper use of Rt, particularly when comparing single drought years, to assess tree resistance to drought. This metric has been much used lately as one of the so-called resilience indices (I wonder why other Rs indices where not included in the manuscript). In the paper, the authors use Rt similarly to most studies in the literature, which is fine. Yet, Rt (and particularly the other resilience indices based on radial growth) does not only depend on the individual response to drought events, but also on stand dynamics around target trees. For this reason, if stands around trees are not comparable or show changes after some disturbance (e.g. the drought events analysed) differently, this may bias comparisons of Resilience indices (although it is true that Rt should be less influenced than the other indices). The problem is that there is not such stand competition data for old droughts. This does not invalidate the study but I think the authors should discuss the issue, particularly for the analysis and comparisons of single droughts. I would suggest including also other Resilience indices, not only Rt, to enrich the discussion.

[*issue #92*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/92)

[*issue #93*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/93)

@@

- Statistical analysis: the analysis would be more precise if the authors used Generalized linear mixed models to avoid data transformation and take into account (with random effects, as they do for the linear mixed model, LMM, presented now) the influence of species in the analysis. Then I missed a further discussion on the role of individual species (e.g. see my comment below on Rt>1).

[*issue #94*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/94)

@@

I suggest to show some relationship, e.g. between Rt and traits in figures. The study only presents two figures and there is plenty of room to present results more clearly with extra figures.

[*issue #71*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/71)

@@

Only assessing models with AIC is not enough to prove sound relationships in multidimensional and complex data-models. LL ratio tests or deviance-anova tests can be used in the case of nested models with LMM or GLMM respectively. Along the text: dAIC is generally referred as $\Delta$AIC, please change accordingly.

[*issue #94*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/94)

*dAIC: minor fix*

@@

Following the two previous comments, I suggest to leave in Table 5 just the best model (rather that a multimodel inference as is presented now) to discuss more clearly the covariates included. And I suggest discussing only the model with all years together. This would be more robust and avoid ramble differences between years that mostly look spurious and with a difficult physiological explanation. This would help to ensure a better use of the Resilience indices.

*This is an interesting suggestion. Depending on where we end up going with some other issues, it may make sense to de-emphasize the individual models--perhaps move Table 5 to SI and replace with figure ([issue #71](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/71))? However, I think we could also just make an argument for leaving this as is.*

related to [*issue #71*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/71)

@@

The authors might like to consider including Table 4 as a supplement, because as it stands it looks like an exploratory analysis. If the ‘single-variable approach’ is not just exploratory, state it clearly and why it is needed besides the full models with several parameters.

[*issue #74*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/74)

@@

- Importantly, the drought selection would be biased if years were selected as from line 202, which is a common mistake in similar studies (the original of Lloret et al. included). Droughts must be selected based only on some climatic/drought index (PDSI, SPEI,…) and independently of growth. Otherwise the analysis is biased towards years with low growth, and different questions would arise with a difficult answer. If the authors select only those droughts where there is a growth decrease: is that decrease evident in all species and similar across species? What happens with more intense droughts that those selected (if any), why they do not force a decrease in growth if drought is a main forcing agent? I see that 1966 and 1999 are the driest (as in PDSI) in Figure 1, but there is at least one year around 1955 that it is as dry or drier than 1977. In line 204 it is stated that they are the driest. If this is the case, the analysis should be sound but it should be deleted what is stated in lines 202-203 where authors say that they selected years also based on low growth. In fact, that should be a result after having selected the drought years only based on dryness (as in lines 208-209). Line 211 then they search for pointer years… a little bit messy. Please, explain better, and it should follow: 1) selection of dry years without regarding growth data; 2) characterise growth those dry years; 3) If the authors want to compare with other pointer years, fine, but that should be a third independent step properly justified and explained. Following this rationale, then in line 221 it should be difficult that Rt>1, had we selected those years with a negative growth anomaly. Please, explain clearly and consistently.

*Ugghh, this doesn't even change anything--we'd still be analyzing the same 3 droughts! The question is whether to remove the growth reduction criteria. I personally think that analyzing 2 criteria is a strength, as our goal was to identify droughts that were both climatologically and biologically signficant. In any case, this is just an issue of changing the wording.*

@@

The issue of the preferential death of large trees is not only important because of its effects on forest function and culture. It is important because it seems to require novel theory. Under current expectations, the largest individuals, with their greater root reach, etc., should be more rather than less resistant. That they might not be demands a novel explanatory framework. Yet the very existence of the tendency is debated, with some studies reporting greater mortality among small rather than large individuals (Camarero et al., 2015; Colangelo et al., 2017). These studies, however, were not designed to take into account the microsite variation that presumably should lead to marked differences in survivorship even within the same height class with drought. One approach, which we have used (Soriano et al., in press; Olson et al., 2018) is to conduct experiments under controlled conditions to detect the effect of height, and our results are consistent with the greater vulnerability of taller individuals. However, under natural conditions, a study that takes this variation into account is needed to test the notion that taller individuals are preferentially vulnerable is needed, and this study provides exactly such a design.

I found this paper not only to address a very important point but also to be exceptionally well structured and written. My main suggestion is that it could use a bit more emphasis on its importance, and I would be interested in hearing a bit more about what the authors feel their results imply regarding the cause of the preferential vulnerability of large individuals. I give line-by-line suggestions (and sometimes just reactions):

*What a nice review! We can easily add some language on the importance.*

@@

102 great questions, though within-species comparisons are likely to be most informative here, and your mixed model design is in accord with this; just being cautious because sometimes it’s seductive to regard selection acting between rather than within species

105 this sounds great, but was a little unclear. When you say “first, we test…” I thought that you would examine a relationship between microsite and some tree characteristic, e.g. that the largest trees are in moist microsites. But the next sentence didn’t refer to a tree-microsite correlation but vulnerability-size. It would be good to clarify the prose here because these are very important passages.  The third sentence did seem to be about tree-microsite correlations.

112 wasn’t clear where the “ring porous more drought resistant” hypothesis came from. Global data show that ring porous species are almost entirely restricted to the north temperate zone, especially deciduous broadleaf forests (Wheeler et al., 2007). They are virtually absent in drylands, especially warm tropical drylands. Species in drylands are almost overwhelmingly diffuse porous; the high wood density species in chaparrals and in some desert plants can be semi ring porous (Olson et al., in press).

*good point. hypothesis refers to temperate deciduous biome.*

@@

113 These correlations seem reasonable, e.g. see (Méndez-Alonzo et al., 2012)

124 and Table 3. This seems like a good palette of species because they are very different in their hydraulic construction. In addition to libriform fibers, oaks have very abundant vasicentric tracheids, especially in latewood. This means that the oaks will likely maintain some conductance even under highly negative tensions. This is not by virtue of their ring porosity (note that most oaks globally aren’t ring porous!), but because of the high proportion of vasicentric tracheids in their latewood. Liriodendron (diffuse porous, scalariform perforation plates) and Carya (likely ring porous at your locality; simple perforation plates) both have libriform fibers and lack vasicentric tracheids. See the data in (Olson et al., in press). There is relatively low family-level diversity represented here; our phylogenetic analyses (Olson et al., in press) suggest that there is very little phylogenetic signal in the conduit diameter-stem length relationship, so this could help justify if necessary not including phylogeny in the analyses.

*this is new terminology to me. Look into Olson et al. in press. Would these wood traits be worth including as covariates?*

@@

L139-140: makes sense to not core live trees in the plot (surprised you got permission to do this in 2011!). But doesn’t this introduce a bias in your analysis? Trees that died may have had very different responses to drought that those which survived. Do you get similar results if you analyse trees that lived and died separately?

L165-167: I think this needs to be explained more clearly. Why are type I errors unlikely and type II error possible? I also think there are real issue with the analysis of crown position. Presumably on average the crown illumination index of trees would have increased over the past 50 or so years (from suppressed to dominant) as trees grew and neighbouring trees died. So I’m not convinced you can include this in your models going back half a century. It’s just not sensible to assume that crown position would not have changed over such a long time period.

L217-226: this is problematic because drought resistance (Rt) as defined by Lloret et al. (2011) assumes that tree growth (BAI in this case) fluctuates around a long term mean that is stable through time. But this is rarely/never the case with tree growth, which shows clear temporal trends related to tree size/ontogeny. If you don’t account for these directional trends then your estimates of Rt will be biased. For instance, if a tree is on a long-term trajectory of decreasing BAI over time, when you calculate Rt you will overestimate the impact of the drought (and vice versa for trees on an upward growth trajectory). So you need to first detrend the tree ring series to remove long-term growth trends (unless you can show that your trees exhibited no long-term temporal trends in BAI). The fact that pre-drought conditions were different for the three droughts is also a problem, but not much you can do about this except acknowledge as you have done.

[*issue #92*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/92)

### 2020-05-17 manuscript/response_to_reviews.Rmd

*Ugghh, this doesn't even change anything--we'd still be analyzing the same 3 droughts! The question is whether to remove the growth reduction criteria. I personally think that analyzing 2 criteria is a strength, as our goal was to identify droughts that were both climatologically and biologically signficant. In any case, this is just an issue of changing the wording.*

**IM - I kind of understand what they meant, but more in terms of a logic flow position. I agree that this can pretty much be fixed**

@@

[*issue #93*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/93)

Secondly, the choice of traits does not appear well justified, as it seems to be driven more by easiness of measurement than by relevance, and it mixes different types of hypotheses operating at different levels. On one hand there is tree height and crown position, which are measured at the level of individual trees and correspond to closely related hypotheses. On the other hand there are ring-porosity, wood density, LMA and water relations traits, which are measured at the species level and seem to be totally disconnected from the hypotheses regarding tree height. In addition, it is very hard to compare the explanatory power of these variables, as done in the paper, as they are based on very different sample sizes (571 trees vs 12 species) and respond to different sources of variability. In fact the way height effects are analyzed mixes the within- and between-species levels, which makes it difficult to interpret the results. Finally, I think it would be more correct to refer to PLA and TLP as water relations traits rather than hydraulic traits.

Thirdly, a more thorough analysis of the distribution of the study species in the sample plot would be required. The paper assumes that all species were subjected to the same drought levels. However, this is unlikely to be true, as different species are likely to occupy different microhabitats in the plot, and hence be subjected to different drought conditions. The authors account for this, to some extent, by including individual-level TWI in some models, but a more thorough analysis of the distribution of the study species as a function of TWI and other relevant micro-environmental variables would be needed to fully interpret the results.

**It seems from these comments they think this shouldn't be published at all due to its limited scope and not enough explanatory power. The resulting study that they suggest, if completely thorough, would take a long time to create. I'm not sure how to address this comment, other than to make a note in the manuscript acknowledging that a more thorough study would be great but this is what we found with what we had.**

### 2020-05-19 manuscript/response_to_reviews.Rmd

## Referee: 1

Comments to the Author

The study addresses the relationship between drought resistance (Rt) calculated from radial growth, and different traits and microenvironmental characteristics on 12 tree species from one forest in the Eastern US. Both the dataset and the study are meaningful, and the relationship found between Rt and height of biological importance. However, I think there are a number of issues that need to be worked out.

It is important to better discuss the proper use of Rt, particularly when comparing single drought years, to assess tree resistance to drought. This metric has been much used lately as one of the so-called resilience indices (I wonder why other Rs indices where not included in the manuscript). In the paper, the authors use Rt similarly to most studies in the literature, which is fine. Yet, Rt (and particularly the other resilience indices based on radial growth) does not only depend on the individual response to drought events, but also on stand dynamics around target trees. For this reason, if stands around trees are not comparable or show changes after some disturbance (e.g. the drought events analysed) differently, this may bias comparisons of Resilience indices (although it is true that Rt should be less influenced than the other indices). The problem is that there is not such stand competition data for old droughts. This does not invalidate the study but I think the authors should discuss the issue, particularly for the analysis and comparisons of single droughts. I would suggest including also other Resilience indices, not only Rt, to enrich the discussion.

@@

[*issue #92*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/92)

**Regarding the resilience indices, we considered analyzing these and decided it would be best to stick with just resistance for three reasons: (1) Most importantly, we would expect resilience to be governed by fundamentally different mechanisms than resistance. While it would be interesting to examine resilience, this would be beyond the scope of the current paper. (2) [DeSoto et al. (2020, Nature)](https://www.nature.com/articles/s41467-020-14300-5) used an extensive dataset of trees sampled after severe droughts and compared the responses to previous, less-severe droughts between trees that survived and trees that were killed by the severe drought. In angiosperms, trees that died from a severe drought had lower resistance to previous droughts, compared to trees that survived the severe drought. In gymnosperms, trees that died during severe droughts had lower recovery following previous droughts, compared to trees that survived the severe drought. These findings would argue that resistance is a more important variable for angiosperms. (3) As the reviewer notes, resilience indices would tend to be more sensitive to a tree's neighborhood and prior conditions. We now state the logic for focusing on $Rt$ in the methods section:**

@@

[*issue #93*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/93)

Statistical analysis: the analysis would be more precise if the authors used Generalized linear mixed models to avoid data transformation and take into account (with random effects, as they do for the linear mixed model, LMM, presented now) the influence of species in the analysis. Then I missed a further discussion on the role of individual species (e.g. see my comment below on Rt>1).

@@

related to [*issue #71*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/71)

The authors might like to consider including Table 4 as a supplement, because as it stands it looks like an exploratory analysis. If the ‘single-variable approach’ is not just exploratory, state it clearly and why it is needed besides the full models with several parameters.

[*issue #74*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/74)

Importantly, the drought selection would be biased if years were selected as from line 202, which is a common mistake in similar studies (the original of Lloret et al. included). Droughts must be selected based only on some climatic/drought index (PDSI, SPEI,…) and independently of growth. Otherwise the analysis is biased towards years with low growth, and different questions would arise with a difficult answer. If the authors select only those droughts where there is a growth decrease: is that decrease evident in all species and similar across species? What happens with more intense droughts that those selected (if any), why they do not force a decrease in growth if drought is a main forcing agent? I see that 1966 and 1999 are the driest (as in PDSI) in Figure 1, but there is at least one year around 1955 that it is as dry or drier than 1977. In line 204 it is stated that they are the driest. If this is the case, the analysis should be sound but it should be deleted what is stated in lines 202-203 where authors say that they selected years also based on low growth. In fact, that should be a result after having selected the drought years only based on dryness (as in lines 208-209). Line 211 then they search for pointer years… a little bit messy. Please, explain better, and it should follow: 1) selection of dry years without regarding growth data; 2) characterise growth those dry years; 3) If the authors want to compare with other pointer years, fine, but that should be a third independent step properly justified and explained. Following this rationale, then in line 221 it should be difficult that Rt>1, had we selected those years with a negative growth anomaly. Please, explain clearly and consistently.

@@

*Ugghh, this doesn't even change anything--we'd still be analyzing the same 3 droughts! The question is whether to remove the growth reduction criteria. I personally think that analyzing 2 criteria is a strength, as our goal was to identify droughts that were both climatologically and biologically signficant. In any case, this is just an issue of changing the wording.*

*IM - I kind of understand what they meant, but more in terms of a logic flow position. I agree that this can pretty much be fixed*

@@

[*issue #93*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/93)

Secondly, the choice of traits does not appear well justified, as it seems to be driven more by easiness of measurement than by relevance, and it mixes different types of hypotheses operating at different levels. On one hand there is tree height and crown position, which are measured at the level of individual trees and correspond to closely related hypotheses. On the other hand there are ring-porosity, wood density, LMA and water relations traits, which are measured at the species level and seem to be totally disconnected from the hypotheses regarding tree height. In addition, it is very hard to compare the explanatory power of these variables, as done in the paper, as they are based on very different sample sizes (571 trees vs 12 species) and respond to different sources of variability. In fact the way height effects are analyzed mixes the within- and between-species levels, which makes it difficult to interpret the results. Finally, I think it would be more correct to refer to PLA and TLP as water relations traits rather than hydraulic traits.

Thirdly, a more thorough analysis of the distribution of the study species in the sample plot would be required. The paper assumes that all species were subjected to the same drought levels. However, this is unlikely to be true, as different species are likely to occupy different microhabitats in the plot, and hence be subjected to different drought conditions. The authors account for this, to some extent, by including individual-level TWI in some models, but a more thorough analysis of the distribution of the study species as a function of TWI and other relevant micro-environmental variables would be needed to fully interpret the results.

*It seems from these comments they think this shouldn't be published at all due to its limited scope and not enough explanatory power. The resulting study that they suggest, if completely thorough, would take a long time to create. I'm not sure how to address this comment, other than to make a note in the manuscript acknowledging that a more thorough study would be great but this is what we found with what we had.*

*I agree.... I wouldn't change analysis for this. We can come back to wording a response later.*
